---
title: "Viz & EDA"
author: "Yiying Wu"
date: "2023-09-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=TRUE,message = FALSE}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

working with NOAA weather data, which is downloaded using `rnoaa::meteo_pull_monitors` function in the code chunk below; similar code underlies the weather dataset used elsewhere in the course. 
```{r,include=TRUE,message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_df
```

## Basic scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
Pipes and stuff
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_nyc_weather=
  weather_df |>
    filter(name=="CerntralPark_NY")|>
    ggplot(aes(x = tmin, y = tmax)) + 
    geom_point()
```

## Fancy Plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```
`geom_smooth`: add a smooth curve

`alpha = .5`:make the data points a bit transparent
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```


`facet_grid(.~name)`: Plot with facets

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth() +
  facet_grid(.~name)
```
Let's try a different plot.

`size = prcp`: higher `prcp` will have bigger points
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

try assigning a specific color
```{r}
weather_df |>
  filter(name=="CentralPark_NY") |>
  ggplot(aes(x=date,y=tmax))+
  geom_point(color="blue") 

weather_df |>
  filter(name!="CentralPark_NY") |>
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point(alpha=.7,size=.5) 
```

`geom_hex`: add density -> the plot will not be messy with too many points
```{r}
weather_df|>
  ggplot(aes(x=tmin,y=tmax))+
  geom_hex()
```

## univaria plotting

histogram
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax,color=name)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax,fill=name)) + 
  geom_histogram()

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

density plots 

`adjust` in density plots is similar to the binwidth parameter in histograms, and it helps to try a few values. 

`alpha`: the transparency level
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = 2, color = "blue")
```

Boxplots

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_boxplot()
```

violin plots
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin()

ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

ridge plot
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges()
```

```{r}
weather_df|>
  filter(name=="CentralPark_NY")|>
  ggplot(aes(x = date, y = tmax))+
  geom_line(alpha=.5)+
  geom_point(size=.5)
```

## Saving and embedding plots
```{r}
ggp_weather = 
  ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) 

ggsave("results/ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```


```{r, fig.width=6,fig.asp=.6}
ggp_weather
```

including the following in a code snippet at the outset of my R Markdown documents.
```
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

