---
title: "Viz & EDA"
author: "Yiying Wu"
date: "2023-09-28"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=TRUE,message = FALSE}
library(tidyverse)
library(ggridges)
```

working with NOAA weather data, which is downloaded using `rnoaa::meteo_pull_monitors` function in the code chunk below; similar code underlies the weather dataset used elsewhere in the course. 
```{r,include=TRUE,message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
weather_df
```

## Basic scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```
Pipes and stuff
```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()

ggp_nyc_weather=
  weather_df |>
    filter(name=="CerntralPark_NY")|>
    ggplot(aes(x = tmin, y = tmax)) + 
    geom_point()
```

## Fancy Plot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```
`geom_smooth`: add a smooth curve

`alpha = .5`:make the data points a bit transparent
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```


`facet_grid(.~name)`: Plot with facets

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color = name)) + 
  geom_point(alpha = .3) +
  geom_smooth() +
  facet_grid(.~name)
```
Let's try a different plot.

`size = prcp`: higher `prcp` will have bigger points
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

try assigning a specific color
```{r}
weather_df |>
  filter(name=="CentralPark_NY") |>
  ggplot(aes(x=date,y=tmax))+
  geom_point(color="blue") 

weather_df |>
  filter(name!="CentralPark_NY") |>
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point(alpha=.7,size=.5) 
```



